MSB1015_Assignment 2

### 2019-09-25 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
Query is changed to only include alkanes without complex structures (so only containing C's and H's). This is an edit from egonw's code. 

Added code to obtain structure descriptors.

---
title: "MSB1015_Assignment 2"
author: "Suus ten Hage"
date: "25 september 2019"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
**Project Description**
In this project I will train a Partial Least Square (PLS) algorithm to predict the bioling point of alkines from their structure. This is based on data from Wiener, 1947.

** 0. Load packages**
If you haven't installed the packages on your computer yet, please uncomment the 
install.packages lines.

```{r}
# install.packages("WikidataQueryServiceR")
library("WikidataQueryServiceR")
```

** 1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to obtain all alkines with known boiling points from WikiData. To do this, I first select all chemical compounds (Q11173), their labels and their boiling points (P2102). Next, I select all compounds with a label that is in English. Finally, I select the compounds whose label ends with "ane" to select the alkanes. This results in 91 alkanes and their boiling points.

```{r}
data = query_wikidata('SELECT DISTINCT ?alkanes ?alkaneLabel ?boilingPoint      WHERE {
    ?alkanes (wdt:P31|wdt:P279) wd:Q11173;
      rdfs:label ?alkaneLabel;
      wdt:P2102 ?boilingPoint.
    FILTER((LANG(?alkaneLabel)) = "en")
    FILTER(STRENDS(?alkaneLabel, "ane"))
  }'
)
```

>>>>

---
title: "MSB1015_Assignment 2"
author: "Suus ten Hage"
date: "25 september 2019"
output: html_notebook
---

**Project Description**
In this project I will train a Partial Least Square (PLS) algorithm to predict the bioling point of alkines from their structure. This is based on data from Wiener, 1947.

**0. Load packages**
If you haven't installed the packages on your computer yet, please uncomment the 
install.packages lines.

```{r}
# install.packages("WikidataQueryServiceR")
library("WikidataQueryServiceR")
#install.packages('rJava')
library('rJava')
#install.packages('rcdk')
library('rcdk')
#install.packages('pls')
library(pls)

```

**1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to obtain all alkines with known boiling points from WikiData. To do this, I first select all chemical compounds (Q11173), their labels and their boiling points (P2102). Next, I select all compounds with a label that is in English and whose label ends with "ane" to select the alkanes. Finally, we select for those alkanes that also have their smile available in the WikiData database. This results in 89 alkanes and their boiling points and smiles.

```{r}
data = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
    p:P2102 [ps:P2102 ?bp ;           
            psv:P2102/wikibase:quantityUnit  
            ?bpUnit         
            ] ;
    wdt:P233 ?smiles.
    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } 
    } '
)

```

**3.Quantify the alkane structures**

```{r}

#Get descriptors
parsed_smiles <- parse.smiles(data$smiles, kekulise=TRUE)

descNames <- c(
'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor'
)

descriptors <- eval.desc(parsed_smiles, descNames)
```
